CF Registry Analysis

### 2023-11-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd
started survival analysis

```{r}
# Import data and codebook
annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
# Analysis function
log_reg <- function(year, missingness = F) {
  # Clean data
  df <- annual %>%
    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
    select(
      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
      A_supplement_feeding, tube_feeding,
      NumPulmExacerbation, A_pseudomonasaeruginosa,
      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
      liver_disease
    )
  # Remove all empty columns
  empty <- which(colSums(is.na(df)) == nrow(df))
  if (length(empty) > 0) {
    df <- df[, -empty]
  }
  # Models
  vars <- colnames(df)[which(!colnames(df) %in%
    c(
      bugs, "cfrd_status_annual", "eDWID", "A_bmivalue",
      "A_bmipercentile", "A_height", "A_heightpercentile",
      "Age_YrEnd", "A_weight", "bmi", "age_perc",
      "A_supplement_feeding", "A_feeding2", "A_feeding3",
      "A_feeding4", "A_FEV1", liver_fields
    ))]
  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
  # Labels
  labels <- list(
    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
    A_FEV1_pct_predicted = "FEV1 percent predicted",
    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
    A_corticosteroids1 = "Use of corticosteroids",
    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
    height_perc = "Height percentile", bmi_perc = "BMI percentile",
    A_Vx770 = "Use of CFTR modulator Vx770",
    A_VX809comb = "Use of CFTR modulator Vx809 combination",
    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
    A_fungalyeast1 = "Aspergillus (any species)",
    A_pulmonarycomplications1 = "ABPA",
    A_mycocultureresults3 = "Microorganisms in myco culture"
  )
  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
    as.character(labels[vars[which(vars %in% colnames(df))]])
  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
    as.character(labels[bugs[which(bugs %in% colnames(df))]])
  # Visualize missingness if required
  if (missingness) {
    cat("\n")
    cat("### Missingness")
    cat("\n")
    cat("\n")
    naniar::vis_miss(df)
    cat("\n")
  }
  # Simple models
  mod1 <- glm(base_mod, data = df, family = "binomial")
  # Print results
  cat("\n")
  cat("### Base Model")
  cat("\n")
  cat("\n")
  print(forest_model(mod1, exclude_infinite_cis = F))
  cat("\n")
  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
    cat("\n")
    cat("### P. aeruginosa")
    cat("\n")
    cat("\n")
    print(forest_model(mod2, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_fungalyeast1" %in% colnames(df)) {
    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
    cat("\n")
    cat("### Aspergillus")
    cat("\n")
    cat("\n")
    print(forest_model(mod3, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_pulmonarycomplications1" %in% colnames(df)) {
    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
    cat("\n")
    cat("### ABPA")
    cat("\n")
    cat("\n")
    print(forest_model(mod4, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_mycocultureresults3" %in% colnames(df)) {
    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
    cat("\n")
    cat("### Mycobacterium")
    cat("\n")
    cat("\n")
    print(forest_model(mod5, exclude_infinite_cis = F))
    cat("\n")
  }
}
```

@@

@@

@@

```{r results='asis',fig.height=10,fig.width=10}
log_reg(2004,missingness = T)
```

>>>>

```{r}
# Import data and codebook
annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
```

@@

Why are the percentages so low in 2006 - 2009? These appear to be correct based on the data we have, so could it be to do with how the data were collected or pulled by the CFFPR?

# CFRD incidence and prevalence

Based on CFFPR annualized data.

```{r}
# Find approximate birth date
annual$approx_birth_year = annual$ReviewYear - ceiling(annual$Age_YrEnd)
annual = annual %>% group_by(eDWID) %>% 
  mutate(approx_birth_year = round(mean(approx_birth_year)))
# Create survival dataset
s <- annual %>%
  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
  group_by(eDWID) %>%
  mutate(start = lag(end)) %>% # Get interval start
  select(eDWID, start,end, Age_YrEnd, approx_birth_year,cfrd_status_annual) # Sort columns
s$start[is.na(s$start)] = s$approx_birth_year[is.na(s$start)]
# Age group as time-varying covariate
s$age_group <- cut(s$Age_YrEnd,
  breaks = c(-Inf, 10, 19, 31, Inf), right = F,
  labels = c("< 10", "[10,19)", "[19,31)", ">= 31")
)
# Fill down diagnosis
s$cfrd_status_annual[s$cfrd_status_annual=="CFRD-"] = NA
s = s %>% group_by(eDWID) %>% fill(cfrd_status_annual,.direction = "down")
s$cfrd_status_annual[is.na(s$cfrd_status_annual)] = "CFRD-"
# Remove records where start == end
s = s[s$end > s$start,]
# Survival fit
surv_obj = Surv(time = as.numeric(s$start),time2 = as.numeric(s$end),
                event = as.numeric(s$cfrd_status_annual)-1)
fit <- survfit(surv_obj ~ age_group,id = eDWID,data = s)
ggsurvplot(fit,risk.table = T,cumevents = T,
  cumcensor = T,
           fun = "event",xlim = c(2003, 2018))
```

@@

```{r}
# Analysis function
log_reg <- function(year, missingness = T) {
  # Clean data
  df <- annual %>%
    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
    select(
      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes, A_FEV1_pct_predicted,
      tube_feeding, NumPulmExacerbation, A_pseudomonasaeruginosa,
      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
      liver_disease
    )
  # Remove all empty columns
  empty <- which(colSums(is.na(df)) == nrow(df))
  if (length(empty) > 0) {
    df <- df[, -empty]
  }
  # Models
  vars <- colnames(df)[which(!colnames(df) %in%
    c(bugs, liver_fields, "cfrd_status_annual", "eDWID"))]
  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
  # Labels
  labels <- list(
    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
    A_FEV1_pct_predicted = "FEV1 percent predicted",
    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
    A_corticosteroids1 = "Use of corticosteroids",
    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
    height_perc = "Height percentile", bmi_perc = "BMI percentile",
    A_Vx770 = "Use of CFTR modulator Vx770",
    A_VX809comb = "Use of CFTR modulator Vx809 combination",
    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
    A_fungalyeast1 = "Aspergillus (any species)",
    A_pulmonarycomplications1 = "ABPA",
    A_mycocultureresults3 = "Microorganisms in myco culture"
  )
  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
    as.character(labels[vars[which(vars %in% colnames(df))]])
  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
    as.character(labels[bugs[which(bugs %in% colnames(df))]])
  # Visualize missingness if required
  if (missingness) {
    cat("\n")
    cat("### Missingness")
    cat("\n")
    cat("\n")
    print(naniar::vis_miss(df))
    cat("\n")
  }
  # Simple models
  mod1 <- glm(base_mod, data = df, family = "binomial")
  # Print results
  cat("\n")
  cat("### Base Model")
  cat("\n")
  cat("\n")
  print(forest_model(mod1, exclude_infinite_cis = F))
  cat("\n")
  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
    cat("\n")
    cat("### P. aeruginosa")
    cat("\n")
    cat("\n")
    print(forest_model(mod2, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_fungalyeast1" %in% colnames(df)) {
    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
    cat("\n")
    cat("### Aspergillus")
    cat("\n")
    cat("\n")
    print(forest_model(mod3, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_pulmonarycomplications1" %in% colnames(df)) {
    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
    cat("\n")
    cat("### ABPA")
    cat("\n")
    cat("\n")
    print(forest_model(mod4, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_mycocultureresults3" %in% colnames(df)) {
    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
    cat("\n")
    cat("### Mycobacterium")
    cat("\n")
    cat("\n")
    print(forest_model(mod5, exclude_infinite_cis = F))
    cat("\n")
  }
}
```

@@

```{r results='asis',fig.height=10,fig.width=10}
log_reg(2004)
```
